---
title: "Setting up for use of Condor CHTC"
author: "Frederick Boehm"
date: "9/10/2018"
output: gihtub_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goals

Goal here is to document the setup & organization of files that I need for using Condor CHTC to do my two-dimensional scans as part of the analyses of the chromosome 2 hotspot.

## Setting

Recall that I have downloaded a Keller 2018 .RData file from DataDryad that contains many of their results. The .RData file, "data/Attie_DO378_eQTL_viewer_v1.Rdata" annotates 139 traits as belonging to the chromosome 2 hotspot.

Following Karl's suggestion, I want to look at not only the local trait that Keller identifies as causal for the hotspot, but also several other local-ish traits that have reasonable univariate LOD scores.

## Plan for using Condor

I'll batch the two-dimensional scan jobs by the local trait identity. So, for each local trait that I choose to examine, I'll have 139 jobs - one for each pairing with a hotspot trait.

I'll organize the data files such that each local trait is its own RData file, while the 139 nonlocal traits are all in a single RData file. We can then use the automatically assigned job number to choose a nonlocal trait for each job.

## Preparing RDS files


```{r}
load("../data/Attie_DO378_eQTL_viewer_v1.Rdata")
ls()
library(tidyverse)
```

```{r}
dataset.islet.rnaseq$expr %>%
  dim
colnames(dataset.islet.rnaseq$expr) %>%
  head
```

We need to get the names of the 139 nonlocal traits that map to chromosome 2 hotspot. 

```{r}
dataset.islet.rnaseq$annots %>%
  colnames
dataset.islet.rnaseq$annots$hotspot
dataset.islet.rnaseq$annots %>%
  filter(hotspot == "chr2") -> hotspot_annot_tib
```

```{r}
dataset.islet.rnaseq$expr %>%
  as_tibble() %>%
  select(as.character(hotspot_annot_tib$gene_id)) -> hotspot_expr_tib
```

```{r}
hotspot_expr_tib %>% dim
```

```{r}
hotspot_expr_tib %>%
  mutate(mouse_id = rownames(dataset.islet.rnaseq$expr)) -> hotspot_expr_tib2
```

```{r}
colnames(hotspot_expr_tib2) %>% tail
```

```{r}
hotspot_expr_tib2 %>% dim
```


## Save tibble with expr data and mouse ID numbers in a RDS file

```{r}
saveRDS(hotspot_expr_tib2, file = "../data-to-condor/hotspot_expr_tib2_keller_chr2.rds")
```

## Identify local traits to study

First, we partition the expression data by subsetting on those traits that are located on chromosome 2 (ie, the gene is part of chromosome 2)

```{r}
dataset.islet.rnaseq$annots %>%
  filter(chr == "2")
```

Where, exactly, is the chromosome 2 hotspot within chromosome 2?

```{r}
dataset.islet.rnaseq$lod.peaks %>%
  dim
```
We see, from the above, that, collectively, the expression traits have 39,958 QTL.

We choose only those traits that map to chromosome 2:

```{r}
dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == "2")
```

We now have 3,458 QTL on chromosome 2.

Recall that Keller et al 2018 studied the gene *Hnf4a* as a putative causal gene. The hotspot is centered around 164 Mb on chromosome 2.

We want to see how many local genes have QTL and are positioned at 150 Mb or greater.

To do this, we'll merge the lod.peaks object (after filtering for chr2 QTL) with the annots object. 


```{r}
dataset.islet.rnaseq$lod.peaks %>%
  filter(chrom == "2") %>%
  left_join(dataset.islet.rnaseq$annots, by = c("annot.id" = "gene_id")) %>%
  filter(chr == "2") %>%
  filter(pos > 150) %>%
  arrange(desc(lod)) -> local_qtl_chr2
# chr, here, refers to the position of the transcript's gene, ie, the chromosome on which the gene is located, and is not derived from QTL data
local_qtl_chr2 %>%
  filter(symbol == "Hnf4a")
```

We see that Hnf4a has a lod scores of 49 at 164.02 Mb on chromosome 2. The gene itself starts and stops at 163.51 and 163.57 on chromosome 2.

I don't need to analyze every trait that is local, near the hotspot, and has a reasonable lod score. But, I should choose among them. I want to consider distance from the hotspot and lod peak height when making these choices. Specifically, I want a collection of maybe 15 traits that have reasonable lod scores and are spread from ~150Mb to the end of chromosome 2.

```{r}
local_qtl_chr2 %>%
  filter(lod > 40)
```

If I consider only those traits with lod above 40, I still have 93 traits. Too many. The top three in the table above are positioned at 155, 164 and 177. Perhaps I should use them, then choose a handful of genes that are within, say, 3 Mb of the center of the hotspot (approximately 164 Mb).

```{r}
local_qtl_chr2 %>%
  filter(lod > 40) %>%
  filter(pos > 160, pos < 168)
```


We now have 19 traits that we might consider (above). It might be worthwhile to examine the allele effects plots for the 19 traits before choosing those traits that we'll use in 2-dimensional scans. Alternatively, we could just examine all 19, ie, do 19 x 139 two-dimensional scans (in 19 distinct submit files).

## Getting the expression values for each trait and each mouse 

We'll use the 19 traits above plus the two traits that are slightly more distant from the hotspot, ie , *Ahcy* and *GM4723*. This gives us a total of 21 local traits.

Rather than making a separate RDS file for each local trait, I'll make one RDS file that contains all of them. I'll then just need to specify, in my condor submit file, which of the traits I want to use in a given condor submission.

```{r}
trait_names1 <- local_qtl_chr2 %>%
  filter(lod > 40) %>%
  filter(pos > 160, pos < 168) %>%
  select(annot.id) %>%
  unlist() %>%
  as.character()
trait_names <- c(trait_names1, "ENSMUSG00000027597", "ENSMUSG00000078898") # add ids for Ahcy and GM4723

```

Now, we subset the expression data matrix to get only the 21 traits that we seek.

```{r}
dataset.islet.rnaseq$expr %>%
  as_tibble() %>%
  select(trait_names) %>%
  mutate(mouse_id = rownames(dataset.islet.rnaseq$expr)) %>%
  saveRDS("../data-to-condor/keller2018-chr2-local-expr.rds")
```


## File transfer to condor

On September 10, 2018, about 10:10 AM, I transferred via scp the two rds files (created above) to the CHTC.

Before doing that, I cloned the git repository (which is also on github) "keller2018-chr2-hotspot-chtc" to my chtc disk space. Specifically, I put it under my home directory, ie, I didn't put it under the existing "attie" directory within my home directory on chtc.






## References

Keller et al. 2018. Genetics. Genetic drivers of pancreatic islet function.

