---
title: "Data Downloading"
author: "Frederick Boehm"
date: "8/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We start with materials that Keller et al. (2018, GENETICS) first reported. The text of their article is available here: 

http://www.genetics.org.ezproxy.library.wisc.edu/content/209/1/335

or, for those not at the U. of Wisconsin, 

http://www.genetics.org/content/209/1/335



We download their data from Data Dryad: 

https://datadryad.org/resource/doi:10.5061/dryad.pj105

Note that the above url gives information on citing and using the data that they share freely. **I'll need to reference this information in my thesis.**

## Reading the data

README.README has info on the objects that are contained within the file "Attie_DO378_eQTL_viewer_v1.Rdata".

Let's load the Rdata file into an R session.

```{r}
load("../data/Attie_DO378_eQTL_viewer_v1.Rdata")
```


## Keller et al 2018, supplementary table 1

Keller et al, in their supplemental materials on figshare, tabulate the 5 hotspots and list the number of traits that demonstrate evidence for mediation at each. They also list a candidate mediator at each hotspot. They also list the number of eQTL for each hotspot, not just the number of genes that are mediated.

This table provides a foundation for generating hypotheses that my test of pleiotropy v separate QTL can address. For example, we see that, on chromosome 2 at 165.5 Mb, 147 expression traits associate. Keller et al. say that the candidate mediator is Hnf4a, and that 88 of the 147 associated expression traits have a large LOD drop in the mediation test. It might be cool to see how many of the 147 expression traits are pleiotropic with the Hnf4a gene expression trait. That is, I could run 147 bivariate tests, each including the Hnf4a gene expression trait and exactly one of the other 147 traits.


I need to see if Keller et al. share a list of the identities of the genes that associate with each hotspot, as well as a list of the genes with mediation test results.

Another question that I have is "How did Keller et al decide on the putative mediator for each of the five hotspots?"

Presumably they describe this in the article text.

